I don't know, are you sure we want to do this?\\ Sea level adaptation decisions under uncertainty

### 2016-12-29 submission/paper.tex
more results on Bergen

We illustrate the sea level rise projections and timing of adaptation measures for two case studies in Northern Europe, namely the city of Bergen in Norway and Esbjerg in Denmark. The Norwegian city of Bergen is the capital of Hordaland County. The city center is located on Byfjorden, and is surrounded by mountains. It has the largest port in Norway, both in terms of freight and passengers. The historic harbor area, Bryggen, is the only Hanseatic trade center remaining in its original style, and has been declared a UN-ESCO World Heritage site\footnote{See \url{http://whc.unesco.org/en/list/59}.}. Bryggen is regularly flooded at extreme tides, and it is feared that as sea levels rise, floods will become a major problem in other parts of Bergen as well \citep{bergenreport}.

@@

We project local sea level changes by Modeling two processes:
\begin{enumerate}
\item The relationship between global temperature and global sea level;
\item The relationship between global sea level and local sea level.
\end{enumerate}

@@

To assess the optimal timing of an adaptation measure, we employ a probabilistic extension of the framework described by \citep{Fankhauser&1999} in which we obtain a probabilistic distribution for the net present value damage in a given year for various adaptation options. The probabilistic distribution is constructed by considering uncertainty
\begin{itemize}
\item in the local sea level projections;
\item in the annual damage costs;
\item in the effect of changes in sea level on the annual damage costs. 
\end{itemize}

@@

\subsection{Limitations of the decision framework}
The main limitation of this light touch decision framework is that we have significantly simplified the assessment of the effect of sea level rise on the damage costs. In particular, the linear extrapolation of the results reported in \citet{Hallegatte&2013} might provide a conservative estimate of the effect of extreme sea level rise. However, with only two data points, extrapolation approaces such as a power law or exponential growth seem unfeasible. Alternatively, a modeling framework similar to that of \citet{Hallegatte&2013} could be applied directly to a larger range of potential changes in sea level. Additionally, we have not specifically accounted for potential changes in storm surge patterns.

@@

\subsection{Timing of adaptation measures {\color{blue} (KdB, TT)}}

The histogram of the observed annual damages and the estimated distribution are given in Figure~\ref{fig:BergenDamageDist}. The parameter estimates for the Burr distribution are $\hat{\alpha} = 1.27$, $\hat{\gamma} = 0.51$ and $\hat{\theta} = 0.002$.

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{DecisionAnalysisBergen.pdf}
\caption{Damage data used in the decision analysis for Bergen: (a) Estimated distribution of annual storm surge damage in Bergen for 2015 (red) based on observed annual damage in Hordaland and Rogaland 1980-2015 (gray bars); (b) Relative change in mean annual damage as a function of sea level rise for 15 European cites as estimated by \cite{Hallegatte&2013} (black circles) with linearly extrapolated values indicated by gray lines. The median change and the corresponding extrapolation are indicated in red.}
\label{fig:BergenDamageDist}
\end{center}
\end{figure}

\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=0.5\linewidth]{CumDamageCostsBergen.pdf}
\caption{Distribution of accumulated future loss 2016-2100 under RCP 8.5 and assuming that no adaptation measures are implemented.} 
\label{fig:NoAction}
\end{center}
\end{figure}

>>>>

\begin{figure}[!hbpt]
\begin{center}
  \includegraphics[width=0.45\linewidth]{BergenMap.png}
  \includegraphics[width=0.45\linewidth]{EsbjergMap.png}
\caption{Terrain maps of central Bergen, Norway (left) and Esbjerg, Denmark (right).}
\label{fig:CityMaps}
\end{center}
\end{figure}

We illustrate the sea level rise projections and timing of adaptation measures for two case studies in Northern Europe, namely the city of Bergen in Norway and Esbjerg in Denmark, see Figure~\ref{fig:CityMaps}. The Norwegian city of Bergen is the capital of Hordaland County. The city center is located on Byfjorden, and is surrounded by mountains. It has the largest port in Norway, both in terms of freight and passengers. The historic harbor area, Bryggen, is the only Hanseatic trade center remaining in its original style, and has been declared a UN-ESCO World Heritage site\footnote{See \url{http://whc.unesco.org/en/list/59}.}. Bryggen is regularly flooded at extreme tides, and it is feared that as sea levels rise, floods will become a major problem in other parts of Bergen as well \citep{bergenreport}.

@@

We project local sea level changes by Modeling two processes, the relationship between global temperature and global sea level, and the relationship between global sea level and local sea level.

@@

To assess the optimal timing of an adaptation measure, we employ a probabilistic extension of the framework described by \citep{Fankhauser&1999} in which we obtain a probabilistic distribution for the net present value damage in a given year for various adaptation options. The probabilistic distribution is constructed by considering uncertainty in the local sea level projections, in the annual damage costs, and in the effect of changes in sea level on the annual damage costs.

@@

\subsection{Limitations of the decision framework}
The main limitation of this light touch decision framework is that we have significantly simplified the assessment of the effect of sea level rise on the damage costs. In particular, the linear extrapolation of the results reported in \citet{Hallegatte&2013} might provide a conservative estimate of the effect of extreme sea level rise. However, with only two data points, extrapolation approaces such as a power law or exponential growth seem unfeasible. Alternatively, a modeling framework similar to that of \citet{Hallegatte&2013} could be applied directly to a larger range of potential changes in sea level. Our framework also simplifies the cost and effect of an adaptation option during construction in that we assume no effect until the construction is finished with all the construction cost falling in the last year of the construction. Especially for larger constructions, these assumptions might need to be modified. Additionally, we have not specifically accounted for potential changes in storm surge patterns.

@@

\subsection{Timing of adaptation measures in Bergen}

\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=0.5\linewidth]{DecisionAnalysisBergenPartI.pdf}
\caption{Estimated distribution of annual damage costs in Bergen for 2015 (red) based on observed annual damage in Hordaland and Rogaland counties 1980-2015 (gray bars).}
\label{fig:BergenDamageDist}
\end{center}
\end{figure}

Figure~\ref{fig:BergenDamageDist} shows the histogram of observed annual damage costs for Bergen and the associated Burr distribution. The parameter estimates are $\hat{\alpha} = 1.27$, $\hat{\gamma} = 0.51$ and $\hat{\theta} = 0.002$. \citet{bergenreport} discuss several different adaptation options for Bergen. In Figure~\ref{fig:TotalDamageBergen} we consider the optimal timing of an adaptation option that includes two inner barriers at V{\aa}gen and Damg{\aa}rdssundet, that is, one on each side of central Bergen. The combined construction cost of the two barriers for a protection against 75 cm sea level rise is estimated at 1.13 billion NOK (2015 level).

\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=\linewidth]{TotalDamageCostsAdaptation.pdf}
\caption{Projected total damage costs in Bergen for the time period 2016-2100 as a function of the timing of an adaptation measure consisting of the construction of two inner barriers. The median projection under each adaptation scenario is indicated in red with gray bars denoting the 90\% projection intervals. The median projected total damage cost under no action is shown with a black line with the corresponding 90\% projection interval indicated by dotted lines. }  
\label{fig:TotalDamageBergen}
\end{center}
\end{figure}

\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=0.5\linewidth]{CumDamageCostsBergen.pdf}
\caption{Median projected cumulative damage costs in Bergen under constant sea level (gray line), under sea level rise according to RCP 8.5 with no adaptation action (black line) and with the construction of two inner barriers in 2047 (red line). The shaded gray area denotes the 90\% projection interval under constant sea level. Dotted lines indicate the 90\% projection intervals with sea level rise according to RCP 8.5. }
\label{fig:NoAction}
\end{center}
\end{figure}
